---
title: "Visualizing Potholes in Boston Streets"
author: "HP Vertica Place Development Team"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    highlight: tango
  html_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    highlight: tango
    theme: journal
---

The `Boston_Pothole_Git.sql` file used Vertica Place to determine the locations of the potholes reported and addressed by the city of Boston, with a focus on determining the intersection of each point location of a pothole and the region in which the pothole occurs. Work through the Place analysis in order to examine how Place formats the data. 

Here, we bring the data into R for visualization. We will use the `ggmap` and `ggplot2` packages in order to create spatial maps of pothole occurrences, using a color gradient to examine the number of potholes per US Census Block.

Code chunks displayed below can be run in your local R instance, provided you've downloaded the corresponding data available in the Vertica/Place Github repository. 


### Loading required libraries
```{r eval=TRUE, echo=TRUE}
library(ggplot2)
library(ggmap)
```

### Importing data output from Place
```{r eval=TRUE, echo=TRUE}
new_potholes = read.csv("heatmap.dat")
```

Next, we'll need to use `ggmap` to obtain a map of downtown Boston and the greater Boston metropolitan area. We then plot pothole occurrences on top of these maps.


```{r eval=TRUE, echo=TRUE,  comment="", message=FALSE}
# Return the maps for different locations
metro <- get_map(location = "Boston, MA", zoom = 11, maptype='toner-hybrid')
downtown <- get_map(location = "Boston, MA", zoom = 15, maptype='toner-hybrid')

# Create the heatmap
heatMapDowntown <- ggmap(downtown) + 
 
  geom_polygon(data = new_potholes, aes(y = Longitude, x = Latitude, 
                                        group = gid, fill = Ratio), alpha = .7) + 
  
  scale_fill_gradient(low = "light yellow", high = "red", guide = "legend", 
                      na.value="light yellow", limits = c(0,5), breaks = c(0:5))


heatMapMetro <- ggmap(metro) + 
  
  geom_polygon(data = new_potholes, aes(y = Longitude, x = Latitude, 
                                        group = gid, fill = Ratio), alpha = .7) + 
  
  scale_fill_gradient(low = "light yellow", high = "red", guide = "legend", 
                      na.value="light yellow", limits = c(0,5), breaks = c(0:5))

```


```{r eval=TRUE, echo=F, tidy=TRUE, comment="", message=FALSE, fig.cap="The spatial distribution of the number of potholes per US Census Block for the greater Boston area. "}
heatMapMetro
```




```{r eval=TRUE, echo=F, tidy=TRUE, comment="", message=FALSE, fig.cap = "The spatial distribution of the number of potholes per US Census Block for downtown Boston."}
heatMapDowntown
```









