---
title: "Cambridge, MA Expired Meters: Jan. 2014 - July 2015"
author: "HPE Vertica Place Development Team"
date: "May 10, 2016"
output:
  html_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    highlight: tango
    theme: journal
---

The `camb_parking.sql` file used Vertica Place to export a shapefile containing the total number of tickets per parking space, the parking meter id, and the parking meter's coordinates.

Here, we bring the data into R for visualization. We will use the `rgdal` and `leaflet` packages in order to create spatial maps of parking spaces, using a color gradient to examine the number of expired meter parking tickets per City of Cambridge metered parking spaces.

Code chunks displayed below can be run in your local R instance, provided you've downloaded the corresponding data available in the Vertica/Place Github repository. 

### Loading required libraries
```{r message=FALSE, warning=FALSE}
library(rgdal)
library(leaflet)
```

### Importing data exported from Place
```{r eval=TRUE, echo=TRUE}
expired_meters <- readOGR('C:/camb_parking/expired_meters/expired_meters.shp',
                          'expired_meters', stringsAsFactors = FALSE)
```

### Creating our color palette

We will use 9 bins. This will give us a range of approximately 48 tickets per bin.

```{r eval=TRUE, echo=TRUE}
binpal <- colorBin("YlOrRd", expired_meters@data$num_ticket, 9, pretty = FALSE)
colors_meters <- binpal(expired_meters@data$num_ticket)
```

### Popups

Here, we'll create popups that display the metered parking space number and the total number of tickets written for that space.

```{r eval=TRUE, echo=TRUE}
popups_expired <- apply(expired_meters@data, 1, function(info) {
  paste0('Expired Meter<br>',
         '<b>Space Number: </b>', info['meter'], '<br>',
         '<b>Number of Tickets: </b>', info['num_ticket'], '<br>')
})
```

### Plot the map

Next, we'll use `leaflet` to obtain an interactive map. Then, we will plot the parking space polygons on top of the map. When you click on a shape, a popup displays the metered parking space number and the total number of tickets written for that space.

```{r eval=TRUE, echo=TRUE}
base_map <- leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addProviderTiles("Stamen.TerrainBackground",
                   options = providerTileOptions(opacity = 0.55)) %>%
  addProviderTiles("Stamen.TonerLabels")

base_map %>% 
  addPolygons(data = expired_meters, 
              color = colors_meters, fillColor=colors_meters,
              popup = unname(popups_expired),
              options = c(pathOptions(), popupOptions(minWidth=100)),
              group = "Expired Meters") %>%
  addLegend(position = "topright", pal = binpal,
            values = expired_meters@data$num_ticket,
            title = "Meters: # of Tickets")
```